---
title: "Fit Series with Litterfitter"
output: html_document
date: "2024-07-08"
---

This markdown fits various analytical decomposition models to time series of leaf litter mass using the package litterfitter (https://github.com/traitecoevo/litterfitter). The initial ranges of model parameters given in litterfitter are appropriate assuming the time variable is measured in years and decomposition progresses relatively slowly compared to the rates typical of aquatic decomposition. However, they've been hand-tuned to encapsulate the ranges necessary to fit the litter data here: I expanded sets of model parameters so that fitted parameters were not clumped on the minimum or maximum for each parameter.

Cornwell, W. K., & Weedon, J. T. (2014). Decomposition trajectories of diverse litter types: a model selection analysis. Methods in Ecology and Evolution 5(2):173-182.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(broom)
library(litterfitter)
library(foreach); library(doSNOW); library(parallel)
library(here)
```

# Data ingest

Data series were checked and  tidied in "tidy_detrital_nutrients_data.Rmd"
```{r}

litter_df <- read_csv(here("data/derived_data/tidied_detrital_nutrients_data.csv"))
```

# Fit litter series

```{r}

if(file.exists(here("data/derived_data/DetNut_litter_fits.rds"))){
  
  litter_fits <- readRDS(here("data/derived_data/DetNut_litter_fits.rds"))

}else{

#set up data
nested_data <- litter_df|>
    select(cohort_id, Meas_Day, Mass_prop)|>
    #filter(cohort_id %in% seq(1:38))|> #for testing
    group_by(cohort_id)|> # needs group_by()|>nest() structure... nest_by() doesn't work!
    nest()|>
  ungroup()|>
  mutate(batch_number = row_number())

# Definitions for use in parallel loops
n_data_batches <- nrow(nested_data)
cohort_names <- nested_data|>pull(cohort_id)
n_iters <- 9999

pb <- txtProgressBar(max = n_data_batches, style = 3)
progress <- function(n) setTxtProgressBar(pb, n)
opts <- list(progress = progress)

# Set parallel processing
cl <- makePSOCKcluster(16) 
registerDoSNOW(cl)


# Loop
litter_fits <- foreach(i = 1:nrow(nested_data), .combine = 'rbind', .options.snow = opts, 
        .packages = c("tidyverse", "litterfitter", "future"), .inorder = TRUE#, .final = function(x) setNames(x, cohort_names)
        ) %dopar% {
          
          # filter to cohort data and unnest into normal data frame
      df <- nested_data|>filter(batch_number == i)|>unnest()
        
       cohort_id <- df|>first()|>pull(cohort_id)
          #negative exponential fit
          negexp_fit <- fit_litter(time = df$Meas_Day,  
                                              mass.remaining = df$Mass_prop,
                                              upper = c(2), 
                                              lower = c(1e-08),
                                              model = "neg.exp",
                                              iters = n_iters/10) # need far fewer iterations for only one parameter
          
          #weibull fit
          weibull_fit <- fit_litter(time = df$Meas_Day, 
                                              mass.remaining = df$Mass_prop,
                                              lower = c(0.1,0.001),#beta, then alpha 
                                              upper = c(25000,100),#beta, then alpha 
                                              model = "weibull",
                                              iters = n_iters)
        
          #discrete parallel fit
          discpar_fit <- fit_litter(time = df$Meas_Day, 
                                              mass.remaining = df$Mass_prop,
                                              upper = c(1, 50, 50), #A, K1, K2 
                                              lower = c(0, 1e-12, 1e-17), #A, K1, K2
                                              model = "discrete.parallel",
                                              iters = n_iters)
        
          #discrete series fit
           discser_fit <- fit_litter(time = df$Meas_Day, 
                                              mass.remaining = df$Mass_prop,
                                              upper = c(1, 50, 2), #R, K1, K2
                                              lower = c(0, 1e-04, 1e-20), #R, K1, K2
                                              model = "discrete.series",
                                              iters = n_iters)
           
           # continuous quality/continuous exponential fit
            contqual_fit = fit_litter(time = df$Meas_Day, 
                                              mass.remaining = df$Mass_prop,
                                              upper = c(10^6, 50), #b,a
                                              lower = c(1e-04, 0.01), #b,a
                                              model = "cont.quality",
                                              iters = n_iters)
            
            # Not actually sure what the name of this is, but appears to estimate a lower mass limit at which decomposition ceases
            negexplim_fit = fit_litter(time = df$Meas_Day, 
                                              mass.remaining = df$Mass_prop,
                                              upper = c(10, 50, 1),
                                              lower = c(1e-06, 1e-02, 1e-20),
                                              model = "neg.exp.limit",
                                              iters = n_iters)
          
            # Package up results for export
            tibble(cohort_id = cohort_id,
                   negexp_fit = list(negexp_fit), 
                    weibull_fit = list(weibull_fit),
                    discpar_fit = list(discpar_fit),
                    discser_fit = list(discser_fit),
                    contqual_fit = list(contqual_fit),
                    negexplim_fit = list(negexplim_fit))
            
   
        }     
stopCluster(cl)

# save to file
saveRDS(litter_fits, file = here("data/derived_data/DetNut_litter_fits.rds")) 

# Loops took about 50 minutes with 16 cores
}
```

