---
title: "Toy Model"
author: "W. J. Matthaeus"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: yes
    mode: selfcontained
  revealjs::revealjs_presentation:
    theme: sky
    highlight: pygments
    center: true
---
Based on the "toy model" from A. Lecerf 2021 section 19.5

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pkgs <- c("cowplot","parallel","doSNOW","tictoc","ggsci","here","readr", "ggplot2", "dplyr", "readxl", 
          "tibble", "tidyr", "stringr", "RColorBrewer",
          "deSolve", "purrr", "forcats")
vapply(pkgs, library, logical(1), character.only = TRUE, logical.return = TRUE)

# Load in negexp fits on real time series
detnut_params <- readRDS(here("data/derived_data/DetNut_litter_fits.rds"))|>
  mutate(negexp_k = map_dbl(negexp_fit, ~pluck(., "optimFit", "par", 1)),
         weibull_alpha = map_dbl(weibull_fit, ~pluck(., "optimFit", "par", 2)))|>
  select(cohort_id, negexp_k,weibull_alpha)
         
```



# Simualate decomp
## ODEs
```{r functions}

# actual model function
toy<-function(t, state, parameters) { 
  with(
    as.list(c(state, parameters)),
    {
     # rate of change 
    dM_l <- -k_1*M_l
    M <- M_l + M_n
    dM_n <- -(4*b*B_max)/(M_0^2) * ((M_0 - M)^(1+c)) * M
    # return the rate of change 
    list(c(dM_l, dM_n)) 
    }
    
    ) 
}

# function to loop across parameter sets
wiggle_ode <- function(initial_mass, proportion_leachable, k_1, b, c, B_max){

  #define variables
  proportion_leachable = proportion_leachable
  initial_mass = initial_mass
  
  temp_params <- c(M_0 = initial_mass,
                proportion_leachable = proportion_leachable,
                k_1 = k_1,
                B_max = B_max,
                b = b,
                c = c
              )
  
    state <- c(M_l = initial_mass*proportion_leachable,
          M_n = initial_mass*(1-proportion_leachable))


# run ODEs
  temp.out <- ode(y = state, times = times, func = toy, parms = temp_params, method = "rk4")
  
#tidy output  
  temp.out.df <- as.data.frame(temp.out)|>
    mutate(M = M_l+M_n)|> 
    filter(time == round(time)) # reduce data points to daily rather than 10ths of a day 
#WJM 1 Aug 24: also can just run a coarser time series. might change the results a little, but it will be faster
}

```

```{r define parameter sets}


times <- seq(0, 1000, by = .1)


params_table <- expand_grid(
  initial_mass = 7.2, # NOTE: for some reason this breaks with a few integer values (e.g., 6) for initial_mass, at least when c != 0 or 1
  M_0 = initial_mass,
  proportion_leachable = c(0.05,0.15,0.25),
  k_1 = c(0.6),
  B_max = c(0.2), 
  b = seq(0.1,0.7, length.out = 5),# must be positive
  #b = seq(0.05,0.45, by = 0.1), #old, wasn't covering high range of 
  c = c(-1,0,0.5)) #lower limit at -1, 1 produces lagging curves that seem implausible - Weibull alphas fairly high


```

```{r run ODEs}


outputs_df_nested <- params_table|>
  rowwise()|>
  mutate(output = pmap(list(initial_mass,proportion_leachable, k_1, b, c, B_max), 
                       ~wiggle_ode(initial_mass = ..1, proportion_leachable = ..2, k_1 = ..3, 
                                   b =..4, c = ..5, B_max = ..6)))|> # tack results onto parameter table
  ungroup()|>
  mutate(param_set_id = row_number())

outputs_df <- outputs_df_nested|>
  unnest(output)|>
  mutate(across(-c(M, M_l, M_n, time), as_factor))|>
  group_by(param_set_id)|>
  mutate(mass_prop = M/first(M),
         mass_prop = round(mass_prop, digits = 2))|>relocate(mass_prop,M) #helpful for some plotting

blah <- outputs_df|>
  filter(M>=0)|>
  #filter(mass_prop)
  arrange(desc(M))|>
  slice_tail(n = 1)
```

## Plots
```{r plot simulations}
ggplot(outputs_df|>filter(mass_prop>0.15))+
  geom_line(aes(x = time, y = M, color = proportion_leachable, linetype = b))+
  facet_wrap(.~c)

ggplot(outputs_df|>filter(mass_prop>0.15))+
  geom_line(aes(x = time, y = log(M), color = proportion_leachable, linetype = b))+
  facet_wrap(.~c)


ggplot(outputs_df|>filter(mass_prop>0.15))+
  geom_line(aes(x = time, y = M, group = param_set_id, color = as_factor(k_1)))

ggplot(outputs_df|>filter(mass_prop>0.15))+
  geom_line(aes(x = time, y = log(M), group = param_set_id, color = as_factor(k_1)))

ggplot(outputs_df|>filter(mass_prop>0.15))+
  geom_line(aes(x = time, y = M, group = param_set_id, color = as_factor(c), linetype = as_factor(proportion_leachable)))+
  facet_wrap(.~as_factor(b))

ggplot(outputs_df|>filter(mass_prop>0.15))+
  geom_line(aes(x = time, y = log(M), group = param_set_id, color = as_factor(c), linetype = as_factor(proportion_leachable)))+
  facet_wrap(.~as_factor(b))

ggplot(outputs_df|>filter(mass_prop>0.15))+
  geom_line(aes(x = time, y = M_n, group = param_set_id, color = as_factor(c), linetype = as_factor(proportion_leachable)))+
  facet_wrap(.~as_factor(b))+
  ggtitle("non-leachable mass fraction only")

ggplot(outputs_df|>filter(mass_prop>0.15))+
  geom_line(aes(x = time, y = log(M_n), group = param_set_id, color = as_factor(c), linetype = as_factor(proportion_leachable)))+
  facet_wrap(.~as_factor(b))+
  ggtitle("non-leachable mass fraction only")

leaching.p <- outputs_df|>
  filter(mass_prop>0.15, time <10)|>
  mutate(proportion_leachable = fct_relevel(as_factor(proportion_leachable),"0.25","0.15","0.05"))|>
  group_by(param_set_id)|>
  mutate(M_l_prop = M_l/first(M_l))|> #plotting this verifies proportionally equivalent (k)
ggplot()+
  geom_line(aes(x = time, y = M_l, linetype = proportion_leachable, color = proportion_leachable), linewidth = 1.25)+
  ylab("Leachable mass")+
  xlab("Time (d)")+
  scale_x_continuous(breaks = c(0,2,4,6,8))+
  scale_y_continuous(breaks = c(0,0.5,1,1.5,2), limits = c(0,2))+
  scale_color_jco()+
  guides(linetype = guide_legend("Leachable Proportion"), color = guide_legend("Leachable Proportion"))+
  theme_classic()+
  theme(axis.title = element_text(face = "bold", size = 18), axis.text = element_text(size = 16, color = "black", face = "bold"),
        legend.title = element_text(size = 14), legend.text = element_text(size = 14), legend.position = c(0.825,0.65))

  ggsave(leaching.p, filename = "leaching_simulation.jpeg", path = here("paper/figures"),
       height = 6, width = 9, units = "in")
  
  
simcurves.p <- outputs_df|>ungroup()|>
  filter(mass_prop>0.15)|>
  mutate(proportion_leachable = fct_relevel(as_factor(proportion_leachable),"0.25","0.15","0.05"),
         c = str_c("c = ", c),
         c = fct_relevel(as_factor(c), "c = -1","c = -0.5", "c = 0", "c = 0.5"))|>
  group_by(param_set_id)|>
  mutate(M_l_prop = M_l/first(M_l))|> #plotting this verifies proportionally equivalent (k)
ggplot()+
  geom_line(aes(x = time, y = M, linetype = proportion_leachable, color = b))+
  facet_wrap(.~c, ncol =2)+
  ylab("Litter Mass (g)")+
  xlab("Time (d)")+
  scale_y_continuous(breaks = c(1, 3,5,7.2), limits = c(1,7.2))+
  scale_x_continuous(limits = c(0,400))+
  scale_color_jco()+
  guides(color = guide_legend("b"), linetype = guide_legend("Leachable Proportion"))+
  theme_classic()+
  theme(axis.title = element_text(face = "bold", size = 18), axis.text = element_text(size = 16, color = "black", face = "bold"),
        legend.title = element_text(size = 14), legend.text = element_text(size = 14), strip.text = element_text(size = 16))
simcurves.p
  ggsave(simcurves.p, filename = "simulated_decomp_curves.jpeg", path = here("paper/figures"),
       height = 6, width = 9, units = "in")
```
k_1 really doesn't influence these curves much at all. Fair to just set k_1 at 0.6 and do leaching scenarios with proportion leachable differences

Interesting that the model  can't ellicit leading dynamics for the non-leachable fraction, where the proportional mass loss (i.e., log(M) ~ time) decreases through time. So, by the model definition, leading dynamics over the whole decomp curve only occur because of leaching. That makes sense with the equation because the per capita decomposer effect (beta) is monotonic. I suppose we could adjust the equation for beta to be something like the probability distribution function for a gamma or weibull distro...

# Analysis
How do negative exponential fits imply biological rates? What is the misrepresentation of k vs a 'correct' biological rate (instantaneous k)? 
```{r calculate weibull and negexp fits on simulated data}
#set up data
nested_outputs <- outputs_df_nested|>
  #slice_head(n = 2)|> # TESTING
  group_by(param_set_id)|>
  mutate(output = map(output, ~.x|>mutate(mass_prop = M/first(M),
                                          mass_prop_nonleachable = M_n/first(M_n))))

if(file.exists(here("data/derived_data/toymodel_fits.rds"))){
  
  sim_output_fits <- readRDS(here("data/derived_data/toymodel_fits.rds"))
  
  }else{


# Definitions for use in parallel loops
n_param_set <- nrow(nested_outputs)
cohort_names <- nested_outputs|>pull(param_set_id)
n_iters <- 999

pb <- txtProgressBar(max = n_param_set, style = 3)
progress <- function(n) setTxtProgressBar(pb, n)
opts <- list(progress = progress)

# Set parallel processing
cl <- makePSOCKcluster(16) 
registerDoSNOW(cl)


# Loop
sim_output_fits <- foreach(i = 1:nrow(nested_outputs), .combine = 'rbind', .options.snow = opts, 
        .packages = c("tidyverse", "litterfitter", "future"), .inorder = TRUE#, .final = function(x) setNames(x, cohort_names)
        ) %dopar% {
    
                
# filter to cohort data and unnest into data frame with full simulation data (all trimmed to end at 0.15 mass proportion, so no super long tails at zero which almost never happens experimentally)
          
      df <- nested_outputs|>filter(param_set_id == i)|>unnest()

# Stop sampling at 0.25 mass prop      
      df_m25 <- df|>filter(mass_prop>=0.25)|> # slightly truncated decomp experiment
        mutate(mass_prop_rounded = round(mass_prop, digits = 2), 
               mass_prop_match = ifelse(mass_prop_rounded %in% round(seq(from = 1-first(proportion_leachable), to = last(mass_prop), length.out = 7), digits = 2),
                                        "yes", "no"))|>
        #filtering to simulate sampling from the time series - sampling starts assuming the first sample timing is just after the majority of the leachate gone
        filter(mass_prop_match == "yes")|>group_by(mass_prop_rounded)|>
        slice_sample(n = 1)

# Stop sampling at 0.45 mass prop 
      df_m45 <- df|>filter(mass_prop>=0.45)|> # mid decomp end
      mutate(mass_prop_rounded = round(mass_prop, digits = 2), 
               mass_prop_match = ifelse(mass_prop_rounded %in% round(seq(from = 1-first(proportion_leachable), to = last(mass_prop), length.out = 7), digits = 2),
                                        "yes", "no"))|>
        #filtering to simulate sampling from the time series - sampling starts assuming the first sample timing is just after the majority of the leachate gone
        filter(mass_prop_match == "yes")|>group_by(mass_prop_rounded)|>
        slice_sample(n = 1)
      

# Stop sampling at 0.6 mass prop      
      df_m60 <- df|>filter(mass_prop>=0.60)|> # short decomposition end
      mutate(mass_prop_rounded = round(mass_prop, digits = 2), 
               mass_prop_match = ifelse(mass_prop_rounded %in% round(seq(from = 1-first(proportion_leachable), to = last(mass_prop), length.out = 7), digits = 2),
                                        "yes", "no"))|>
        #filtering to simulate sampling from the time series - sampling starts assuming the first sample timing is just after the majority of the leachate gone
        filter(mass_prop_match == "yes")|>group_by(mass_prop_rounded)|>
        slice_sample(n = 1)
        
      
 
# Fit simulated data sets and components to models
      
          #negative exponential fit
          negexp_fit <- fit_litter(time = df$time,  
                                              mass.remaining = df$mass_prop,
                                              upper = c(2), 
                                              lower = c(1e-08),
                                              model = "neg.exp",
                                              iters = n_iters) 
          
           #scenario assuming measurement stopped at 0.25*M0
           negexp_fit_m25 <- fit_litter(time = df_m25$time,  
                                              mass.remaining = df_m25$mass_prop,
                                              upper = c(2), 
                                              lower = c(1e-08),
                                              model = "neg.exp",
                                              iters = n_iters) 
           
           #scenario assuming measurement stopped at 0.45*M0
            negexp_fit_m45 <- fit_litter(time = df_m45$time,  
                                              mass.remaining = df_m45$mass_prop,
                                              upper = c(2), 
                                              lower = c(1e-08),
                                              model = "neg.exp",
                                              iters = n_iters) 
            
             #scenario assuming measurement stopped at 0.60*M0
             negexp_fit_m60 <- fit_litter(time = df_m60$time,  
                                              mass.remaining = df_m60$mass_prop,
                                              upper = c(2), 
                                              lower = c(1e-08),
                                              model = "neg.exp",
                                              iters = n_iters) 
          
            negexp_fit_nonleachable <- fit_litter(time = df$time,  
                                              mass.remaining = df$mass_prop_nonleachable,
                                              upper = c(2), 
                                              lower = c(1e-08),
                                              model = "neg.exp",
                                              iters = n_iters) 
            
          
          #weibull fit
          weibull_fit <- fit_litter(time = df$time, 
                                              mass.remaining = df$mass_prop,
                                              lower = c(0.1,0.001),#beta, then alpha 
                                              upper = c(25000,100),#beta, then alpha 
                                              model = "weibull",
                                              iters = n_iters)
          
          weibull_fit_nonleachable <- fit_litter(time = df$time, 
                                              mass.remaining = df$mass_prop_nonleachable,
                                              lower = c(0.1,0.001),#beta, then alpha 
                                              upper = c(25000,100),#beta, then alpha 
                                              model = "weibull",
                                              iters = n_iters)
        
          
            # Package up results for export
            tibble(param_set_id = i,
                   negexp_fit = list(negexp_fit),
                   negexp_fit_m25 = list(negexp_fit_m25),
                   negexp_fit_m45 = list(negexp_fit_m45),
                   negexp_fit_m60 = list(negexp_fit_m60),
                   negexp_fit_nonleachable = list(negexp_fit_nonleachable),
                    weibull_fit = list(weibull_fit),
                   weibull_fit_nonleachable = list(weibull_fit_nonleachable))
            
   
        }     
stopCluster(cl)

saveRDS(sim_output_fits, file = here("data/derived_data/toymodel_fits.rds")) 
}

# Loops took <5 minutes with 16 cores on 999 iters
 
```


```{r}

tidy_outputs <- nested_outputs|>
  left_join(sim_output_fits)|>
  mutate(negexp_k = map_dbl(negexp_fit, ~pluck(., "optimFit", "par", 1)),
         negexp_AICc = map_dbl(negexp_fit, ~pluck(., "fitAICc")),
         negexp_logLik = map_dbl(negexp_fit, ~pluck(., "logLik")),
         negexp_BIC = map_dbl(negexp_fit, ~pluck(., "fitBIC")),
         negexp_pred = map(negexp_fit, ~pluck(., "predicted"))
  )|>
  mutate(negexp_k_m25 = map_dbl(negexp_fit_m25, ~pluck(., "optimFit", "par", 1)),
         negexp_AICc_m25 = map_dbl(negexp_fit_m25, ~pluck(., "fitAICc")),
         negexp_logLik_m25 = map_dbl(negexp_fit_m25, ~pluck(., "logLik")),
         negexp_BIC_m25 = map_dbl(negexp_fit_m25, ~pluck(., "fitBIC")),
         negexp_pred_m25 = map(negexp_fit_m25, ~pluck(., "predicted"))
  )|>
  mutate(negexp_k_m45 = map_dbl(negexp_fit_m45, ~pluck(., "optimFit", "par", 1)),
         negexp_AICc_m45 = map_dbl(negexp_fit_m45, ~pluck(., "fitAICc")),
         negexp_logLik_m45 = map_dbl(negexp_fit_m45, ~pluck(., "logLik")),
         negexp_BIC_m45 = map_dbl(negexp_fit_m45, ~pluck(., "fitBIC")),
         negexp_pred_m45 = map(negexp_fit_m45, ~pluck(., "predicted"))
  )|>
  mutate(negexp_k_m60 = map_dbl(negexp_fit_m60, ~pluck(., "optimFit", "par", 1)),
         negexp_AICc_m60 = map_dbl(negexp_fit_m60, ~pluck(., "fitAICc")),
         negexp_logLik_m60 = map_dbl(negexp_fit_m60, ~pluck(., "logLik")),
         negexp_BIC_m60 = map_dbl(negexp_fit_m60, ~pluck(., "fitBIC")),
         negexp_pred_m60 = map(negexp_fit_m60, ~pluck(., "predicted"))
  )|>
  mutate(weibull_beta = map_dbl(weibull_fit, ~pluck(., "optimFit", "par", 1)),
         weibull_alpha = map_dbl(weibull_fit, ~pluck(., "optimFit", "par", 2)),
         weibull_AICc = map_dbl(weibull_fit, ~pluck(., "fitAICc")),
         weibull_logLik = map_dbl(weibull_fit, ~pluck(., "logLik")),
         weibull_BIC = map_dbl(weibull_fit, ~pluck(., "fitBIC")),
         weibull_pred = map(weibull_fit, ~pluck(., "predicted"))
         )|>
  mutate(negexp_k_nonleachable = map_dbl(negexp_fit_nonleachable, ~pluck(., "optimFit", "par", 1)),
         negexp_AICc_nonleachable = map_dbl(negexp_fit_nonleachable, ~pluck(., "fitAICc")),
         negexp_logLik_nonleachable = map_dbl(negexp_fit_nonleachable, ~pluck(., "logLik")),
         negexp_BIC_nonleachable = map_dbl(negexp_fit_nonleachable, ~pluck(., "fitBIC")),
         negexp_pred_nonleachable = map(negexp_fit_nonleachable, ~pluck(., "predicted"))
  )|>
  mutate(weibull_beta_nonleachable = map_dbl(weibull_fit_nonleachable, ~pluck(., "optimFit", "par", 1)),
         weibull_alpha_nonleachable = map_dbl(weibull_fit_nonleachable, ~pluck(., "optimFit", "par", 2)),
         weibull_AICc_nonleachable = map_dbl(weibull_fit_nonleachable, ~pluck(., "fitAICc")),
         weibull_logLik_nonleachable = map_dbl(weibull_fit_nonleachable, ~pluck(., "logLik")),
         weibull_BIC_nonleachable = map_dbl(weibull_fit_nonleachable, ~pluck(., "fitBIC")),
         weibull_pred_nonleachable = map(weibull_fit_nonleachable, ~pluck(., "predicted"))
         )|>
  select(-contains("fit"), -output, -contains("_pred"))


```



```{r}
tidy_outputs2 <- tidy_outputs|>
  #filter(negexp_k > 0.002)|> # remove some oddly slow ones - keeps close to range in empirical DetNut data, but also published in Follstad-Shah 2017
  mutate(m15_logdiffnl = log(negexp_k)-log(negexp_k_nonleachable),
         m25_logdiffnl = log(negexp_k_m25)-log(negexp_k_nonleachable),
         m45_logdiffnl = log(negexp_k_m45)-log(negexp_k_nonleachable),
         m60_logdiffnl = log(negexp_k_m60)-log(negexp_k_nonleachable),
         m15_diffnl = negexp_k-negexp_k_nonleachable,
         m25_diffnl = negexp_k_m25-negexp_k_nonleachable,
         m45_diffnl = negexp_k_m45-negexp_k_nonleachable,
         m60_diffnl = negexp_k_m60-negexp_k_nonleachable,
         m25_logdiff = log(negexp_k_m25)-log(negexp_k),
         m45_logdiff = log(negexp_k_m45)-log(negexp_k),
         m60_logdiff = log(negexp_k_m60)-log(negexp_k),
         m25_diff = negexp_k_m25-negexp_k,
         m45_diff = negexp_k_m45-negexp_k,
         m60_diff = negexp_k_m60-negexp_k,
         m25_hldiff = 0.693/negexp_k_m25-0.693/negexp_k,
         m45_hldiff = 0.693/negexp_k_m45-0.693/negexp_k,
         m60_hldiff = 0.693/negexp_k_m60-0.693/negexp_k)|>
  mutate(proportion_leachable = as_factor(proportion_leachable*100),
         proportion_leachable = fct_relabel(proportion_leachable, ~str_c(.x, " % Leachable")))


k_diffs_nl <- tidy_outputs2|>
  pivot_longer(cols = c(m15_diffnl,m25_diffnl,m45_diffnl,m60_diffnl), values_to = "k_diff", names_to = "terminal_prop_mass")|>
  relocate(k_diff, terminal_prop_mass)|>
  mutate(terminal_prop_mass = as.numeric(str_remove_all(terminal_prop_mass, "[^0-9]"))/100,
         terminal_prop_mass2 = as_factor(terminal_prop_mass),
         terminal_prop_mass= fct_reorder(terminal_prop_mass2, terminal_prop_mass,.desc = TRUE),
         c = as_factor(c))|>
  select(-terminal_prop_mass2)

k_logdiffs_nl <- tidy_outputs2|>
  pivot_longer(cols = c(m15_logdiffnl, m25_logdiffnl,m45_logdiffnl,m60_logdiffnl), values_to = "k_diff_ln", names_to = "terminal_prop_mass")|>
  relocate(k_diff_ln, terminal_prop_mass)|>
  mutate(terminal_prop_mass = as.numeric(str_remove_all(terminal_prop_mass, "[^0-9]"))/100,
         terminal_prop_mass2 = as_factor(terminal_prop_mass),
         terminal_prop_mass= fct_reorder(terminal_prop_mass2, terminal_prop_mass,.desc = TRUE),
         c = as_factor(c))|>
  select(-terminal_prop_mass2)

k_diffs <- tidy_outputs2|>
  pivot_longer(cols = c(m25_diff,m45_diff,m60_diff), values_to = "k_diff", names_to = "terminal_prop_mass")|>
  relocate(k_diff, terminal_prop_mass)|>
  mutate(terminal_prop_mass = as.numeric(str_remove_all(terminal_prop_mass, "[^0-9]"))/100,
         terminal_prop_mass2 = as_factor(terminal_prop_mass),
         terminal_prop_mass= fct_reorder(terminal_prop_mass2, terminal_prop_mass,.desc = TRUE),
         c = as_factor(c))|>
  select(-terminal_prop_mass2)

hl_diffs <- tidy_outputs2|>
  pivot_longer(cols = c(m25_hldiff,m45_hldiff,m60_hldiff), values_to = "half.life_diff", names_to = "terminal_prop_mass")|>
  relocate(half.life_diff, terminal_prop_mass)|>
  mutate(terminal_prop_mass = as.numeric(str_remove_all(terminal_prop_mass, "[^0-9]"))/100,
         terminal_prop_mass2 = as_factor(terminal_prop_mass),
         terminal_prop_mass= fct_reorder(terminal_prop_mass2, terminal_prop_mass,.desc = TRUE),
         c = as_factor(c))|>
  select(-terminal_prop_mass2)

k_logdiffs <- tidy_outputs2|>
  pivot_longer(cols = c(m25_logdiff,m45_logdiff,m60_logdiff), values_to = "k_diff_ln", names_to = "terminal_prop_mass")|>
  relocate(k_diff_ln, terminal_prop_mass)|>
  mutate(terminal_prop_mass = as.numeric(str_remove_all(terminal_prop_mass, "[^0-9]"))/100,
         terminal_prop_mass2 = as_factor(terminal_prop_mass),
         terminal_prop_mass= fct_reorder(terminal_prop_mass2, terminal_prop_mass,.desc = TRUE),
         c = as_factor(c))|>
  select(-terminal_prop_mass2)


range(tidy_outputs2$negexp_k)
range(detnut_params$negexp_k)

range(tidy_outputs2$weibull_alpha)
range(detnut_params$weibull_alpha)

k_for_plotting <- tidy_outputs2|>ungroup()|>select(weibull_alpha,negexp_k)|>mutate(type = "simulated")|>
  bind_rows(detnut_params|>select(-cohort_id)|>mutate(type = "empirical"))

ggplot(k_for_plotting)+
  geom_density(aes(x = negexp_k, color = type))

ggplot(k_for_plotting)+
  geom_histogram(aes(x = negexp_k, color = type))+
  facet_wrap(.~type)

ggplot(k_for_plotting|>filter(weibull_alpha <30))+
  geom_density(aes(x = weibull_alpha, color = type))

ggplot(k_for_plotting|>filter(weibull_alpha <30))+
  geom_histogram(aes(x = weibull_alpha, color = type))+
  facet_wrap(.~type)
```


```{r difference plots}


klogdiff.p<-k_logdiffs|>
  ggplot()+
  geom_point(aes(x = terminal_prop_mass, y = k_diff_ln, color = c, size = b), alpha = 0.5, position = position_dodge(width = 0.75))+
  scale_size_continuous(range = c(3,6))+
  scale_y_continuous(limits = c(-1,1.25))+
  scale_color_jco(guide = guide_legend(override.aes = list(size = 4)))+
  facet_wrap(.~proportion_leachable)+
  xlab("Mass proportion at terminal measurement")+
  ylab(expression("log(k)"[trunc] - "log(k)"[simulated]))+
  theme_bw()+
  theme(axis.title = element_text(face = "bold", size = 18), axis.text = element_text(size = 16, color = "black", face = "bold"),
        legend.title = element_text(size = 26, face = "bold"), legend.text = element_text(size = 25), strip.text = element_text(size = 18))

klogdiff_nl.p<-k_logdiffs_nl|>
  ggplot()+
  geom_point(aes(x = terminal_prop_mass, y = k_diff_ln, color = c, size = b), alpha = 0.5, position = position_dodge(width = 0.75))+
  scale_size_continuous(range = c(3,6))+
  scale_y_continuous(limits = c(-1,1.25))+
  scale_color_jco(guide = guide_legend(override.aes = list(size = 4)))+
  facet_wrap(.~proportion_leachable)+
  xlab("Mass proportion at terminal measurement")+
  ylab(expression("log(k)"[trunc] - "log(k)"[simulated-nonleachable]))+
  theme_bw()+
 theme(axis.title = element_text(face = "bold", size = 18), axis.text = element_text(size = 16, color = "black", face = "bold"),
        legend.title = element_text(size = 18, face = "bold"), legend.text = element_text(size = 25), strip.text = element_text(size = 18))


kdiff.p<-k_diffs|>
  ggplot()+
  geom_point(aes(x = terminal_prop_mass, y = k_diff, color = c, size = b), alpha = 0.5, position = position_dodge(width = 0.75))+
  scale_size_continuous(range = c(3,6))+
  scale_y_continuous(limits = c(-0.025,0.025))+
  scale_color_jco(guide = guide_legend(override.aes = list(size = 4)))+
  facet_wrap(.~proportion_leachable)+
  xlab("Mass proportion at terminal measurement")+
  ylab(expression("k"[trunc] - "k"[simulated]))+
  theme_bw()+
  theme(axis.title = element_text(face = "bold", size = 18), axis.text = element_text(size = 16, color = "black", face = "bold"),
        legend.title = element_text(size = 26, face = "bold"), legend.text = element_text(size = 25), strip.text = element_text(size = 18))

kdiff_nl.p<-k_diffs_nl|>
  ggplot()+
  geom_point(aes(x = terminal_prop_mass, y = k_diff, color = c, size = b), alpha = 0.5, position = position_dodge(width = 0.75))+
  scale_size_continuous(range = c(1.5,5))+
  scale_y_continuous(limits = c(-0.025,0.025))+
  scale_color_jco()+
  facet_wrap(.~proportion_leachable)+
  xlab("Mass proportion at terminal measurement")+
  ylab(expression("k"[trunc] - "k"[simulated-nonleachable]))+
  theme_bw()+
  theme(axis.title = element_text(face = "bold", size = 18), axis.text = element_text(size = 16, color = "black", face = "bold"),
        legend.title = element_text(size = 18), legend.text = element_text(size = 14), strip.text = element_text(size = 18))

#half-life diffs plot
hldiff.p<-hl_diffs|>
  ggplot()+
  geom_point(aes(x = terminal_prop_mass, y = half.life_diff, color = c, size = b), alpha = 0.5, position = position_dodge(width = 0.75))+
  scale_size_continuous(range = c(3,6))+
  #scale_y_continuous(limits = c(-0.025,0.025))+
  scale_color_jco(guide = guide_legend(override.aes = list(size = 4)))+
  facet_wrap(.~proportion_leachable)+
  xlab("Mass proportion at terminal measurement")+
  ylab(expression("t"[0.5-trunc] - "t"[0.5-simulated] ~(days)))+
  theme_bw()+
  theme(axis.title = element_text(face = "bold", size = 18), axis.text = element_text(size = 16, color = "black", face = "bold"),
        legend.title = element_text(size = 26, face = "bold"), legend.text = element_text(size = 25), strip.text = element_text(size = 18))
hldiff.p

ggplot(hl_diffs,aes(x = half.life_diff))+
  geom_histogram()

```

```{r diff cowplots}

logdiff_grid <- ggdraw(plot_grid(plot_grid(klogdiff.p+theme(legend.position = "none"),
            klogdiff_nl.p+theme(legend.position = "none"), 
            ncol = 1),
          plot_grid(get_legend(klogdiff.p),ncol = 1),
          rel_widths = c(1,0.2)))

ggsave(logdiff_grid, filename = "k_logdiff_sims.jpeg", path = here("paper/figures"),
       height = 9, width = 12, units = "in")

diff_grid <- ggdraw(plot_grid(plot_grid(kdiff.p+theme(legend.position = "none"),
            hldiff.p+theme(legend.position = "none"), 
            ncol = 1),
          plot_grid(get_legend(kdiff.p),ncol = 1),
          rel_widths = c(1,0.2)))

ggsave(diff_grid, filename = "Fig4_diff_sims.jpeg", path = here("paper/figures"),
       height = 9, width = 12, units = "in")

```



  proportion_leachable = c(0.05,0.15,0.25),
  k_1 = c(0.4, 0.8, 1.2),
  B_max = c(0.2,1.2, by = 0.2),
  b = seq(0.2,1, by = 0.2),
  c = seq(-1,2, by = 0.4))

```{r filter down to scenarios}
blah <- tidy_outputs2|>
  mutate(leaching_scenario = case_when(proportion_leachable == 0.25 ~ "high",
                                       proportion_leachable == 0.15 ~ "med",
                                       proportion_leachable == 0.05 ~ "low",
                                       .default = NA_character_),
         decomp_scenario = case_when(b == 0.2 & B_max == 0.2 ~ "low", 
                                     b == 0.6 & B_max == 0.2 ~ "med",
                                     b == 1 ~ "high",
                                     .default = NA_character_))

ggplot(blah|>filter(!is.na(leaching_scenario)&!is.na(decomp_scenario))|>
         filter(k_diff_ln <2))+
  geom_point(aes(x = leaching_scenario, y = k_diff_ln, color = decomp_scenario))
```
